---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Set the number formatter (the number of decimal places)
```{r}
options(scipen=1, digits=2)
```

## Loading and preprocessing the data
```{r}
zipFileName <- "activity.zip"
data <- read.csv(unzip(zipFileName))
```

## Total number of steps taken per day
```{r}
stepsPerDay <- aggregate(data$steps, by=list(data$date), FUN=sum)
colnames(stepsPerDay) <- c("date", "stepsPerDay")
stepsPerDay
```

```{r}
hist(stepsPerDay$stepsPerDay, xlab="Number of steps per day", main="Steps per day")
```

## The mean and median of the total number of steps taken per day
Mean (without NA values): `r mean(stepsPerDay$stepsPerDay, na.rm=TRUE)`
Median (without NA values): `r median(stepsPerDay$stepsPerDay, na.rm=TRUE)`


## What is the average daily activity pattern?

avgDailyPattern <- aggregate(data$steps, by=list(data$interval), FUN=mean, na.action=na.omit)
colnames(avgDailyPattern) <- c("interval", "averageActivity")

## Imputing missing values

I could either remove all NA values or replace them with zeros, because if a person was not using the tracker he or she tracked zero steps.
The following code replaces all NA values with zeros.
```{r}
dataWithoutNa <- data
dataWithoutNa[is.na(data)] <- 0
```

## Are there differences in activity patterns between weekdays and weekends?
